Efficient Planning for Sokoban with Answer Set Programming - A Case Study

### 2018-09-10 BATexFiles/Draft.tex
Cleaning up the encodings and optimizations, not only in the folders but also internally

\chapter{State of the Art}

@@

\textbf{Relevance Cutting:} A basic AI to solve a planning problem like Sokoban will look at all possible ways to move around the level, but some of those are dubious but legal or just plain irrelevant. In a lot of cases those dubious ways do not lead to an optimal solution, but of course a problem can be constructed where the optimal solution can only be found when looking at those dubious cases. So Junghanns and Schaeffer tried to teach their AI meta-reasoning to give it the ability to know when to look at those dubious ways and when to ignore them. This they did by creating an influence measurement and only taking moves that are influenced by prior moves, to prevent actions that are disembodied from all prior moves (and seem irrelevant).

@@

In my research I focussed on simple approaches in very specific domains, to be able to compare the efficiency of different optimizations in those domains. The domains I researched are rectangular halls and slightly more crooked levels, for reference see Figure \ref{fig:domains}. Those maps had to be selfmade, so I had to find the horizon myself (by using incremental encodings). The basic optimizations are some approaches to deadlock detection, basic move-evaluation (don't take moves back), splitting x and y to reduce groundable variables, and i tried for lexordering the ways to reduce multiple grounded ways from A to B.

@@

In the code 

@@

\lstinputlisting{../Encodings/SokobanPlainEncoding.lp}

@@

\subsection{Incremental Encoding}
\paragraph*{•}
\lstinputlisting{../Encodings/SokobanPlainEncodingInc.lp}
The incremental encoding is similar to the plain encoding. The difference is the partition of the encoding into three parts: the \#program base in lines 3-5 which is independent of the step parameter t, the \#program step in lines 7-19 which is the cumulative part, collecting knowledge and the \#program check in lines 21-25 which is specific for each value of t (STEP ALSO IS THOUGH). This allows for gradually processing each time step and accumulating knowledge that persists even if the current step is not satisfiable.

\subsection{Advanced Encoding}
\paragraph*{•}
\lstinputlisting{../Encodings/SokobanPlainEncodingOpti.lp}
\section{Evaluation}
\paragraph*{•}
\section{Experiments and Results}
\paragraph*{•}
\chapter{Discussion of Results}
\paragraph*{•}
compare chapter State of the Art with chapter My Work \cite{LifschitzASP}

>>>>

\chapter{Related Work}

@@

\textbf{Relevance Cutting:} A basic AI to solve a planning problem like Sokoban will look at all possible ways to move around the level, but some of those are dubious but legal or just plain irrelevant. In a lot of cases those dubious ways do not lead to an optimal solution, but of course a problem can be constructed where the optimal solution can only be found when looking at those dubious cases. So Junghanns and Schaeffer tried to teach their AI meta-reasoning to give it the ability to know when to look at those dubious ways and when to ignore them. This they did by creating an influence measurement and only taking moves that are influenced by prior moves, to prevent actions that are disembodied from all prior moves. This is used in the popular Sokoban Solver Rolling Stone which has an excellent record in approaching difficult Sokoban problems.

@@

In my research I focused on simple approaches in very specific domains, to be able to compare the efficiency of different optimizations in those domains. The domains I researched are rectangular halls and slightly more crooked levels, for reference see Figure \ref{fig:domains}. There are few common Sokoban domains that only look at levels with these qualities, so I designed a problemset myself. This meant that I had to find the horizon aswell (by using incremental encodings). The basic optimizations are some approaches to deadlock detection, basic move-evaluation (don't take moves back), splitting x and y to reduce groundable variables, and i tried for lexordering the ways to reduce multiple grounded ways from A to B.

@@

For illustration, consider the instance shown in Figure \ref{fig:levelVis} and the corresponding constants and facts in listing ABC. The turn limit is fixed by the constant \textit{horizon} and the level bounds are fixed by the constants \textit{dimX} and \textit{dimY}. The information about the level is stored in three different atoms:
The \textit{init(field/2)} predicate stores the information about the levelstructure by holding the field coordinates of every field on which the player or a box can be placed.
The \textit{init(at/3)} predicate holds the initial player- and box-positions. The first argument is the identifier of the object. A zero represents the player; all boxes have identifiers greater than zero. The last two arguments are the initial position of the object in the level.
The \textit{init(target/2)} predicate has two arguments in levels containing no assignments. Here the two arguments represent the goal coordinates that have to be covered by a box to satisfy the level. In a level with assignments added there will be another argument in front of the other two referencing the identifier of the box that has to be placed on this goal position.

@@

\lstinputlisting{../EncodingsClean/SokobanPlainEncoding.lp}

@@

\subsection{Incremental Encoding}
\paragraph*{•}
\lstinputlisting{../EncodingsClean/SokobanIncEncoding.lp}
The incremental encoding is similar to the plain encoding. The difference is the partition of the encoding into three parts: the \#program base in lines 3-5 which is independent of the step parameter t, the \#program step in lines 7-19 which is the cumulative part, collecting knowledge and the \#program check in lines 21-25 which is loaded to check for each value of t and then unloaded for the next step. This allows for gradually processing each time step and accumulating knowledge that persists even if the current step is not satisfiable, greatly reducing the time in comparison to reloading the PlainEncoding everytime while incrementing the horizon by one.

\subsection{Advanced Encoding}
\paragraph*{•}
\lstinputlisting{../EncodingsClean/Optimizations/Clutter.optimization.lp}
\section{Evaluation}
\paragraph*{•}
\section{Experiments and Results}
\paragraph*{•}
\chapter{Discussion of Results}
\paragraph*{•}
compare chapter State of the Art with chapter My Work \cite{LifschitzASP}
